---
title: "Untitled"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NorSySS

We observe that NorSySS data during monday to friday is generally lower than the weekend. This implies that analysing the current week will result in the current week being artificially lower.

```{r include=FALSE, results="hide"}
d <- copy(data$covid19$norsyss_norge)
d[,day := lubridate::wday(date, week_start=1)]
d[, day:= factor(
  day,
  levels = c(1:7),
  labels = c("man","tir","ons","tor","fre",glue::glue("l{fhi::nb$oe}r"),glue::glue("s{fhi::nb$oe}n"))
  )]
d[, pr100 := 100*n/consult_with_influenza]
```

Here we can see the results from a negative binomial regression using daily national data (outcome: covid19 consultations, exposure: day of week, offset: total number of consultations)

```{r}
fit <- MASS::glm.nb(n ~ day + offset(log(consult_with_influenza)), data=d)
summary(fit)
```


```{r, echo = F}
q <- ggplot(d, aes(x=day, y=pr100))
q <- q + geom_boxplot()
q <- q + scale_x_discrete(NULL)
q <- q + scale_y_continuous("Andel konsultasjoner",
  breaks = fhiplot::pretty_breaks(5),
  labels = fhiplot::format_nor_perc_0,
  expand = expansion(mult = c(0, 0.1))
)
q <- q + expand_limits(y=0)
q <- q + labs(title = "NorSySS: Andel covid19 konsultasjoner (R991+R992) per dag")
q <- q + labs(caption = glue::glue("Folkehelseinstituttet, {lubridate::today()}"))
q <- q + fhiplot::theme_fhi_lines(
  panel.grid.major.x = element_blank(),
  panel.grid.minor.x = element_blank()
)
q
```

## Alerts

```{r include=FALSE, results="hide"}
d <- copy(data$covid19$norsyss)
setorder(d,location_code,yrwk)
d[,pr100:=100*n/consult_with_influenza]
d[is.nan(pr100), pr100:=0]
d[,yrwk_id := paste0("yrwk",1:.N), by=.(location_code)]

yrwks <- unique(d$yrwk)

d_wide <- dcast.data.table(d, location_code ~ yrwk_id, value.var = c("n","consult_with_influenza","pr100"))

d_wide[,baseline := pmax(0.01,100*(n_yrwk1+n_yrwk2)/(consult_with_influenza_yrwk1+consult_with_influenza_yrwk2))]

d_wide[,threshold_yrwk_1 := 0]
d_wide[,threshold_yrwk_2 := 0]

d_wide[,threshold_yrwk_3 := 100*qpois(0.975, lambda=baseline*consult_with_influenza_yrwk3/100)/consult_with_influenza_yrwk3]
d_wide[is.nan(threshold_yrwk_3),threshold_yrwk_3:=0]

d_wide[,threshold_yrwk_4 := 100*qpois(0.975, lambda=baseline*consult_with_influenza_yrwk4/100)/consult_with_influenza_yrwk4]
d_wide[is.nan(threshold_yrwk_4),threshold_yrwk_4:=0]

d_wide_norsyss <- d_wide

location_code_norsyss <- d_wide_norsyss[pr100_yrwk3 > threshold_yrwk_3 | pr100_yrwk4 > threshold_yrwk_4]$location_code

d <- copy(data$covid19$msis)
setorder(d,location_code,yrwk)
d[,yrwk_id := paste0("yrwk",1:.N), by=.(location_code)]

d_wide <- dcast.data.table(d, location_code ~ yrwk_id, value.var = "n")
d_wide[,baseline := pmax(1,round((yrwk1+yrwk2)/2))]
d_wide[,threshold := qpois(0.975, lambda = baseline)]

d_wide_msis <- d_wide
location_code_msis <- d_wide_msis[yrwk3 > threshold | yrwk4 > threshold]$location_code

location_codes <- unique(c(location_code_norsyss, location_code_msis))

tab_norsyss <- melt.data.table(
  d_wide_norsyss[location_code %in% location_codes],
  id="location_code",
  measure = patterns("^n_", "^pr100_","^threshold_"),
  value.name = c("norsyss_n","norsyss_pr100","norsyss_pr100_threshold")
)

tab_norsyss

tab_msis <- melt.data.table(
  d_wide_msis[location_code %in% location_codes],
  id.vars =c("location_code","threshold"),
  measure.vars = c("yrwk1","yrwk2","yrwk3","yrwk4")
)
levels(tab_msis$variable) <- 1:4

tab <- merge(
  tab_msis,
  tab_norsyss,
  by=c("location_code","variable")
)

setnames(tab,c("threshold","value"),c("msis_threshold","msis_n"))

tab[, pretty_msis_threshold:=fhiplot::format_nor(msis_threshold)]
tab[, pretty_msis_n:=fhiplot::format_nor(msis_n)]
tab[, pretty_norsyss_n:=fhiplot::format_nor(norsyss_n)]
tab[, pretty_norsyss_pr100:=fhiplot::format_nor_perc_1(norsyss_pr100)]
tab[, pretty_norsyss_pr100_threshold:=fhiplot::format_nor_perc_1(norsyss_pr100_threshold)]

tab[variable %in% 1:2, pretty_msis_threshold:=""]
tab[variable %in% 1:2, pretty_norsyss_pr100_threshold:=""]

tab[,location_name := get_location_name(location_code)]
tab

areas_at_risk_ht <- function(tab, yrwks){
  msis_index_hig <- which(tab$msis_n > tab$msis_threshold & tab$variable %in% 3:4)
  norsyss_index_hig <- which(tab$pretty_norsyss_pr100 > tab$pretty_norsyss_pr100_threshold & tab$variable %in% 3:4)
  
  levels(tab$variable) <- yrwks
  
  ht <- huxtable::hux(
    "Geo"=tab$location_name,
    "Uke"=tab$variable,
    "Tilfeller"=tab$pretty_msis_n,
    "Terskel"=tab$pretty_msis_threshold,
    "Konsultasjoner"=tab$pretty_norsyss_n,
    "Andel"=tab$pretty_norsyss_pr100,
    "Terskel"=tab$pretty_norsyss_pr100_threshold
  )%>%
    huxtable::add_colnames() %>%
    fhiplot::huxtable_theme_fhi_basic()
  
  if (length(msis_index_hig) > 0) huxtable::background_color(ht)[msis_index_hig+1, 3] <- fhiplot::warning_color[["hig"]]
  if (length(norsyss_index_hig) > 0) huxtable::background_color(ht)[norsyss_index_hig+1, 6] <- fhiplot::warning_color[["hig"]]
  
  
  ht <- huxtable::add_rows(ht, ht[1, ], after = 0)
  
  ht <- huxtable::merge_cells(ht, 1:2, 1)
  ht <- huxtable::merge_cells(ht, 1:2, 2)
  
  ht <- huxtable::merge_cells(ht, 1, 3:4)
  ht[1, 3] <- "MSIS"
  
  ht <- huxtable::merge_cells(ht, 1, 5:7)
  ht[1, 5] <- "NorSySS"
  
  #huxtable::left_border(ht)[, c(3, 5)] <- 5
  huxtable::left_border_style(ht)[, c(3, 5)] <- "double"
  
  ht <- huxtable::merge_repeated_rows(ht, huxtable::everywhere, 1)
  huxtable::width(ht) <- 1
  huxtable::tabular_environment(ht) <- "longtable"
  ht
}
```

```{r, results="asis", echo=F}
location_names <- unique(tab$location_name)
areas_at_risk_ht(tab[location_name %in% location_names[1:10]], yrwks = yrwks)
areas_at_risk_ht(tab[location_name %in% location_names[11:20]], yrwks = yrwks)
areas_at_risk_ht(tab[location_name %in% location_names[21:30]], yrwks = yrwks)
areas_at_risk_ht(tab[location_name %in% location_names[31:40]], yrwks = yrwks)
areas_at_risk_ht(tab[location_name %in% location_names[41:42]], yrwks = yrwks)
```
